---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, echo = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "README-"
)
```

# concrete
[![CRAN](http://www.r-pkg.org/badges/version/concrete)](http://www.r-pkg.org/pkg/concrete)
[![R-CMD-check](https://github.com/imbroglio-dc/concrete/actions/workflows/R-CMD-check.yaml/badge.svg)](https://github.com/imbroglio-dc/concrete/actions/workflows/R-CMD-check.yaml)
[![Coverage Status](https://coveralls.io/repos/github/imbroglio-dc/concrete/badge.svg?branch=main)](https://coveralls.io/github/imbroglio-dc/concrete?branch=main)

> Continuous-Time Targeted Maximum Likelihood Estimation for Survival Analysis with Competing Risks

## Description

`concrete` is an R package for estimating covariate-adjusted marginal cumulative incidence (absolute risk) in right-censored survival settings with and without competing risks. It implements the continuous-time one-step Targeted Minimum Loss-Based Estimation (TMLE) methodology of [Rytgaard et al. (2023)](https://doi.org/10.1111/biom.13856) and [Rytgaard and van der Laan (2023)](https://doi.org/10.1007/s10985-022-09576-2).

TMLE is a doubly-robust, semiparametric efficient estimation framework. `concrete` combines it with cross-validated super learner ensembles for the nuisance parameters (treatment propensity scores and cause-specific hazards), yielding estimates that are:

- **Doubly robust**: consistent if either the propensity score or the outcome model is correctly specified
- **Semiparametrically efficient**: achieves the lowest possible asymptotic variance among regular estimators
- **Targeted**: the TMLE fluctuation step directly reduces bias for the estimand of interest (cumulative incidence)

---

## Installation

Install from [CRAN](https://cran.r-project.org/):

```{r cran-installation, eval = FALSE}
install.packages("concrete")
```

Install the development version from GitHub via [`devtools`](https://www.rstudio.com/products/rpackages/devtools/):

```{r gh-installation, eval = FALSE}
devtools::install_github("imbroglio-dc/concrete")
```

---

## Quick Start

The `concrete` workflow has three steps: format inputs, run TMLE, extract output.

```{r quickstart, eval = FALSE}
library(data.table)
library(concrete)

# --- 1. Prepare data ---
# DataTable must have columns: EventTime, EventType (0 = censored), Treatment, and covariates
data <- as.data.table(survival::pbc)[1:200, .(id, time, status, trt, age, sex)]

# --- 2. Format arguments ---
# makeITT() specifies an intent-to-treat analysis for a binary treatment
concrete.args <- formatArguments(
  DataTable   = data,
  EventTime   = "time",       # observed follow-up time
  EventType   = "status",     # 0 = censored, 1 = transplant, 2 = death
  Treatment   = "trt",        # binary treatment (0/1)
  ID          = "id",
  TargetTime  = c(1000, 2000, 3000),  # time points to estimate risk at
  TargetEvent = c(1, 2),              # competing events to target
  Intervention = makeITT(),           # compare treat-all vs. treat-none
  CVArg       = list(V = 5)           # 5-fold cross-validation
)

# --- 3. Run TMLE ---
concrete.est <- doConcrete(concrete.args)

# --- 4. Extract and visualize results ---
concrete.out <- getOutput(
  concrete.est,
  Estimand     = c("Risk", "RD", "RR"),  # absolute risk, risk diff, risk ratio
  Simultaneous = TRUE                     # simultaneous CIs across time points
)

print(concrete.out)
plot(concrete.out)
```

---

## Core Functions

### `formatArguments()` — Validate and format inputs

`formatArguments()` validates your data and builds a `ConcreteArgs` object consumed by `doConcrete()`. Key arguments:

| Argument | Description |
|---|---|
| `DataTable` | `data.table` with columns for event time, event type, treatment, and covariates |
| `EventTime` | Column name of the observed follow-up time (positive reals) |
| `EventType` | Column name of the event type; values ≤ 0 indicate censoring |
| `Treatment` | Column name of the binary treatment variable (coded 0/1) |
| `TargetTime` | Time point(s) at which to estimate cumulative incidence |
| `TargetEvent` | Event type(s) to target (subset of non-censoring `EventType` values) |
| `Intervention` | List of intervention regimes; use `makeITT()` for intent-to-treat |
| `CVArg` | Arguments for cross-validation folds (passed to `origami::make_folds()`) |
| `Model` | Named list of learners for propensity and hazard models (see below) |
| `GComp` | If `TRUE`, also returns g-computation plug-in estimates for comparison |

**Data requirements:** `DataTable` must contain at minimum `EventTime`, `EventType`, and `Treatment` columns. All remaining columns are treated as baseline covariates.

### `makeITT()` — Specify interventions

`makeITT()` returns a two-element list defining the treat-all (`trt = 1`) and treat-none (`trt = 0`) interventions for a binary, point-in-time treatment. You can also write custom intervention functions directly:

```{r custom-intervention, eval = FALSE}
# Custom intervention: treat if age > 50, otherwise don't treat
my_intervention <- list(
  "high.age" = list(
    intervention = function(trt, W) as.numeric(W[["age"]] > 50),
    gstar        = function(trt, W) as.numeric(W[["age"]] > 50)
  ),
  "low.age" = list(
    intervention = function(trt, W) as.numeric(W[["age"]] <= 50),
    gstar        = function(trt, W) as.numeric(W[["age"]] <= 50)
  )
)
```

### `doConcrete()` — Run the TMLE

`doConcrete()` takes a `ConcreteArgs` object and runs the full estimation pipeline:

1. **Initial estimation**: fits cross-validated propensity scores (via `SuperLearner`) and cause-specific hazards (via Cox regression ensembles)
2. **EIC computation**: computes the efficient influence curve for each target (event, time, intervention) combination
3. **TMLE update**: iteratively fluctuates the hazard estimates along the least favorable submodel until the empirical mean EIC is below the convergence threshold (`|PnEIC| ≤ seEIC / (√n · log n)`)

Returns a `ConcreteEst` object. You can print it to inspect convergence and initial estimator weights, or plot it to view nuisance weight distributions and the convergence trajectory:

```{r plot-est, eval = FALSE}
print(concrete.est)             # convergence status + super learner coefficients
plot(concrete.est, convergence = TRUE, gweights = TRUE)  # diagnostic plots
```

### `getOutput()` — Extract estimates

`getOutput()` extracts final estimates from a `ConcreteEst` object. Supported estimands:

| `Estimand` | Description |
|---|---|
| `"Risk"` | Cause-specific absolute risk (cumulative incidence) under each intervention |
| `"RD"` | Risk difference between two interventions |
| `"RR"` | Risk ratio between two interventions |

```{r getoutput-example, eval = FALSE}
# Absolute risks only
getOutput(concrete.est, Estimand = "Risk")

# Risk difference and ratio (requires exactly 2 interventions)
getOutput(concrete.est, Estimand = c("RD", "RR"), Intervention = c(1, 2))

# Include g-computation plug-in estimates alongside TMLE
getOutput(concrete.est, Estimand = c("Risk", "RD"), GComp = TRUE)

# Simultaneous confidence bands across target times (recommended for multiple comparisons)
getOutput(concrete.est, Estimand = c("Risk", "RD", "RR"), Simultaneous = TRUE)
```

The returned `ConcreteOut` is a `data.table` with columns `Time`, `Event`, `Estimand`, `Intervention`, `Estimator`, `Pt Est`, `se`, `CI Low`, `CI Hi`, and (if `Simultaneous = TRUE`) `SimCI Low`, `SimCI Hi`. Call `plot()` on it to generate `ggplot2` figures.

---

## Model Specification

By default, `concrete` uses a sensible set of learners. You can override this by passing a named list to `Model`:

```{r model-spec, eval = FALSE}
model <- list(
  # SuperLearner library for the propensity score (treatment model)
  "trt" = c("SL.glm", "SL.glmnet", "SL.ranger"),

  # Cox regression formula(s) for censoring hazard (EventType == 0)
  "0" = list(Surv(time, status == 0) ~ age + sex),

  # Cox regression formula(s) for event 1 hazard
  "1" = list(
    "mod1" = Surv(time, status == 1) ~ trt,
    "mod2" = Surv(time, status == 1) ~ .   # "." includes all covariates
  ),

  # Cox regression formula(s) for event 2 hazard
  "2" = list(Surv(time, status == 2) ~ .)
)

concrete.args <- formatArguments(
  DataTable    = data,
  EventTime    = "time",
  EventType    = "status",
  Treatment    = "trt",
  ID           = "id",
  TargetTime   = 2500,
  TargetEvent  = c(1, 2),
  Intervention = makeITT(),
  Model        = model,
  CVArg        = list(V = 5)
)
```

Multiple formulas per event type are combined as a Cox regression ensemble (their linear predictors are stacked and re-fit). The `glmnet` (Coxnet) and `survivalSL` packages are also supported as hazard learners when installed.

A previously generated `ConcreteArgs` object can be passed back into `formatArguments()` to modify individual model components:

```{r modify-args, eval = FALSE}
# Modify the censoring model after the initial formatArguments() call
concrete.args[["Model"]][["0"]] <- list(Surv(time, status == 0) ~ trt + age)
concrete.args <- formatArguments(concrete.args)  # re-validates
```

---

## Output and Inference

Results from `getOutput()` include:

- **Point estimates** (`Pt Est`): marginal cumulative incidence under the specified intervention at each target time
- **Standard errors** (`se`): derived from the efficient influence curve
- **Pointwise CIs** (`CI Low`, `CI Hi`): Wald-type intervals at the specified `Signif` level (default 95%)
- **Simultaneous CIs** (`SimCI Low`, `SimCI Hi`): confidence bands that jointly cover all target time points, computed via multivariate normal simulation from the EIC covariance matrix

```{r output-example, eval = FALSE}
concrete.out <- getOutput(
  concrete.est,
  Estimand     = c("Risk", "RD", "RR"),
  Simultaneous = TRUE,
  Signif       = 0.05   # 95% CI
)

# Tabular output
print(concrete.out)

# ggplot2 figures (one per estimand type)
plot(concrete.out, NullLine = TRUE, ask = FALSE)
```

---

## A note on bootstrap standard errors

A non-parametric bootstrap is **not** offered as an inference option in `concrete`. The reason is covariate imputation: if a user's pipeline involves multiple imputation of missing covariates before calling `concrete`, a correct bootstrap must re-impute covariates on each resample. `concrete` has no visibility into imputation steps that occur upstream, so any bootstrap it ran internally would hold the imputed values fixed — systematically underestimating variance and producing anti-conservative confidence intervals. Users who want bootstrap inference should implement the full resample-then-impute-then-estimate loop themselves, outside the package.

---

## References

Rytgaard, H. C., Gerds, T. A., & van der Laan, M. J. (2023). Continuous-time targeted minimum loss-based estimation of intervention-specific mean outcomes. *Biometrics*, 79(4), 3057–3068. <https://doi.org/10.1111/biom.13856>

Rytgaard, H. C., & van der Laan, M. J. (2023). One-step targeted maximum likelihood estimation for directed acyclic graphs with multiple time point interventions. *Lifetime Data Analysis*, 29, 286–309. <https://doi.org/10.1007/s10985-022-09576-2>

van der Laan, M. J., & Rose, S. (2011). *Targeted Learning: Causal Inference for Observational and Experimental Data*. Springer.

---

## Issues

If you encounter any bugs or have any specific feature requests, please [file an issue](https://github.com/imbroglio-dc/concrete/issues).
