# concrete
[![CRAN](http://www.r-pkg.org/badges/version/concrete)](http://www.r-pkg.org/pkg/concrete)
[![R-CMD-check](https://github.com/imbroglio-dc/concrete/workflows/R-CMD-check/badge.svg)](https://github.com/imbroglio-dc/concrete/actions)
[![Coverage Status](https://coveralls.io/repos/github/imbroglio-dc/concrete/badge.svg?branch=main)](https://coveralls.io/github/imbroglio-dc/concrete?branch=main)

> Continuous-Time Targeted Maximum Likelihood Estimation for Survival Analysis with Competing Risks 

## Description

`concrete` is an R package designed to use targeted maximum likelihood
estimation (TMLE) to compute covariate-adjusted marginal cumulative incidence
estimates in right-censored survival settings with and without competing risks. 

---

## Installation

For standard use, we recommend installing the package from
[CRAN](https://cran.r-project.org/) via

```{r cran-installation, eval = FALSE}
install.packages("concrete")
```

You can install a stable release of `concrete` from GitHub via
[`devtools`](https://www.rstudio.com/products/rpackages/devtools/) with:

```{r gh-installation, eval = FALSE}
devtools::install_github("imbroglio-dc/concrete")
```

---

## Issues

If you encounter any bugs or have any specific feature requests, please [file an
issue](https://github.com/imbroglio-dc/concrete/issues).
