---
title: "concrete vignette using the pbc dataset"
output: html_notebook
---
# Loading data and libraries

```{r libraries}
library(survival, quietly = TRUE)
library(tidyverse, quietly = TRUE)
library(data.table, quietly = TRUE)
library(zoo, quietly = TRUE)
library(nleqslv, quietly = TRUE)
library(sl3, quietly = TRUE)
library(origami, quietly = TRUE)
# devtools::install_github("imbroglio-dc/concrete")
library(concrete, quietly = TRUE)
```

pbc dataset is pulled from the survival package, see ?survival::pbc for details about the dataset. Since we only mean to use this dataset for illustrating the use of `concrete`, we arbitrarily fill in missing treatment assignments below.
```{r pbc data}
set.seed(12345)
data <- as.data.table(survival::pbc)
data[is.na(trt), trt := sample(1:2, sum(is.na(trt)), replace = TRUE)][, trt := trt - 1]
# data[, status := as.numeric(status >= 1)]
data
```

```{r plot pbc}
# plot data
library(prodlim)
for (j in seq_along(unique(data$status)[unique(data$status) > 0])) {
    plot(prodlim(Hist(time, status) ~ strata(trt), data = data),
         cause = unique(data$status)[unique(data$status) > 0][j],
         add = j > 1, lty = j, confint = FALSE, legend = FALSE)
}
```

# concrete
## arguments
### specify target events, times, and interventions
```{r specify targets}
# specify targets -----------------------------------------------------------------------------
target.event <- sort(unique(data[status > 0, status]))
target.time <- quantile(data$time, 2:6/7)

# specify intervention(s) ------------------------------------------------------------------------
# list of arguments to be passed into origami::make_folds(), see ?origami::make_folds for details

# intervention <- list("A == 1" = list("intervention" = function(a, L) {rep_len(1, length(a))},
#                                      "g.star" = function(a, L) {as.numeric(a == 1)}),
#                      "A == 0" = list("intervention" = function(a, L) {rep_len(0, length(a))},
#                                      "g.star" = function(a, L) {as.numeric(a == 0)}))

intervention <- concrete:::IntentToTreat
```

### specify estimation parameters
```{r estimation params}
# cross-validation using origami::make_folds()
cv.arg <- list(n = nrow(data), fold_fun = folds_vfold, cluster_ids = NULL, strata_ids = NULL)

# sl libraries
prop.score.backend <- "sl3"
lrnr_glm <- Lrnr_glm$new()
lrnr_hal <- Lrnr_hal9001$new()
a_lrnrs <- make_learner(Stack, lrnr_glm, lrnr_hal) # sl3 learner object: see tlverse sl3 documentation

haz.est.backend <- "coxph" 
cens_lib <- list(model.null = Surv(time, status == 0) ~ 1,
                 Surv(time, status == 0) ~ trt + age + sex)
evnt1_lib <- list(mod.main = Surv(time, status == 1) ~ trt + age + sex,
                  mod.int = Surv(time, status == 1) ~ trt:age + trt:sex)
evnt2_lib <- list(mod1 = Surv(time, status == 2) ~ trt + age + sex)

# concrete model list
model <- list("Trt" = a_lrnrs,
              "0" = cens_lib,
              "1" = evnt1_lib,
              "2" = evnt2_lib)
```

### specify tmle parameters
```{r tmle params}
max.update.iter <- 100
one.step.eps <- 0.1
min.nuisance <- 0.05
```

### Check concrete arguments
```{r concrete arg check}
concrete.args <- formatArguments(Data = data[, c("time", "status", "trt", "id", "age", "sex")],
                                 EventTime = "time", EventType = "status",
                                 Treatment = "trt", ID = "id", Intervention = intervention,
                                 TargetTime = target.time, TargetEvent = target.event,
                                 CVArg = cv.arg, Model = model, PropScoreBackend = prop.score.backend, 
                                 HazEstBackend = haz.est.backend, MaxUpdateIter = max.update.iter, 
                                 OneStepEps = one.step.eps, MinNuisance = min.nuisance, 
                                 GComp = TRUE, Verbose = FALSE)
```

## estimation
```{r, cache=TRUE}
concrete.est <- doConcrete(ConcreteArgs = concrete.args)
```

## output estimates
```{r}
concrete.outputs <- getOutput(Estimate = concrete.est, Estimand = c("RR", "RD", "Risk"),
                              TargetTime = target.time, TargetEvent = target.event, GComp = TRUE)
concrete.outputs$RR
attr(concrete.outputs$RR, "regime")
```

```{r plot-gcomp}
# plot estimates ------------------------------------------------------------------------------
for (j in seq_along(unique(data$status)[unique(data$status) > 0])) {
    plot(prodlim(Hist(time, status) ~ strata(trt), data = data),
         cause = unique(data$status)[unique(data$status) > 0][j],
         add = j > 1, lty = j, confint = FALSE, legend = FALSE)
}
tmp.gcomp <- lapply(concrete.outputs$Risk, function(risks) {
    risks <- setDT(risks)[Estimator == "gcomp", ]
    y <- lapply(unique(risks$Event), function(j) {
        risk.j <- risks[Event == j, ]
        points(risk.j$Time, risk.j$Risk, type = "b")
    })
})
```
```{r plot-tmle}
for (j in seq_along(unique(data$status)[unique(data$status) > 0])) {
    plot(prodlim(Hist(time, status) ~ strata(trt), data = data),
         cause = unique(data$status)[unique(data$status) > 0][j],
         add = j > 1, lty = j, confint = FALSE, legend = FALSE)
}
tmp.tmle <- lapply(concrete.outputs$Risk, function(risks) {
    risks <- setDT(risks)[Estimator == "tmle", ]
    y <- lapply(unique(risks$Event), function(j) {
        risk.j <- risks[Event == j, ]
        points(risk.j$Time, risk.j$Risk, type = "b", pch = 2)
    })
})
```
```{r}

```


